Optimizing a PoS Tag Set for Dependency Parsing

### 2016-05-07 submission.tex
typos

In the current paper, we argue that it is neverthelesss possible to significantly improve the performance
of the downstream application of syntactic parsing by introducing
linguistically informed, syntactically informative distinctions to a tag set.
In the following section we first turn to the treebank that provides the basis for our experiments and
outline its morphological and syntactic annotation choices.

@@

\subsection{Baseline Experiments}
In an initial round of experiments, we concatenated the tag of each token with
its full set of morphological features, thereby mapping the original tag set to a
new maximally fine-grained tag set, given the annotations available in the
treebank. This resulted in a total of 368 tags,
hereafter referred to as the \emph{full} tag set.
The two initial tag sets, i.e., the original tag set comprising 19 tags and the
full tag set comprising 368 tags, thus represent two extremes in terms of
granularity.
To establish some initial points of reference for how tag set granularity affects the
performance of tagging and parsing on NDT, we trained and tested a full pipeline with both of
these intital tags sets. The results are reported in Table \ref{inittagseteval}.
Unsurprisingly, we see that the tagger accuracy
plummets when we move from the original to the full tag set. While the MFT
tagging baseline for the original tag set is 94.14\%, it drops by almost 9
percentage points to 85.15\% for the full tag set. TnT yields 
97.47\% accuracy on the original tag set, which drops to 93.48\% for the full tag
set. These results confirm our hypothesis that the high level linguistic
information
in the full, fine-grained tag set comes at the expense of reduced tagger
performance.  In spite of this, however, we see that the additional information
in the full tag set still improves the parser performance. With the original tag set, Mate
achieves 87.01\% LAS and 90.19\% UAS, which for the full tag set increases to 87.15\% and
90.39\%, respectively. These
preliminary results are encouraging in indicating that the additional linguistic
information assists syntactic parser, motivating further optimization of the
tag set.

>>>>

In the current paper, we argue that it is nevertheless possible to significantly improve the performance
of the downstream application of syntactic parsing by introducing
linguistically informed, syntactically informative distinctions to a tag set.
In the following section we first turn to the treebank that provides the basis for our experiments and
outline its morphological and syntactic annotation choices.

@@

\subsection{Baseline Experiments}
In an initial round of experiments, we concatenated the tag of each token with
its full set of morphological features, thereby mapping the original tag set to a
new maximally fine-grained tag set, given the annotations available in the
treebank. This resulted in a total of 368 tags,
hereafter referred to as the \emph{full} tag set.
The two initial tag sets, i.e., the original tag set comprising 19 tags and the
full tag set comprising 368 tags, thus represent two extremes in terms of
granularity.
To establish some initial points of reference for how tag set granularity affects the
performance of tagging and parsing on NDT, we trained and tested a full pipeline with both of
these initial tags sets. The results are reported in Table \ref{inittagseteval}.
Unsurprisingly, we see that the tagger accuracy
plummets when we move from the original to the full tag set. While the MFT
tagging baseline for the original tag set is 94.14\%, it drops by almost 9
percentage points to 85.15\% for the full tag set. TnT yields 
97.47\% accuracy on the original tag set, which drops to 93.48\% for the full tag
set. These results confirm our hypothesis that the high level linguistic
information
in the full, fine-grained tag set comes at the expense of reduced tagger
performance.  In spite of this, however, we see that the additional information
in the full tag set still improves the parser performance. With the original tag set, Mate
achieves 87.01\% LAS and 90.19\% UAS, which for the full tag set increases to 87.15\% and
90.39\%, respectively. These
preliminary results are encouraging in indicating that the additional linguistic
information assists syntactic parser, motivating further optimization of the
tag set.
